---
title: "About Data"
output: html_document
date: "2024-12-05"
---

```{r, message=FALSE}
library(sf)
library(dplyr)
library(readr)
library(stringr)
library(lubridate)
library(timeDate)
library(suncalc)
library(purrr)
library(readxl)
library(tidyverse)
```

## Data source

### Original data

This report analyzes shooting incidents in NYC. You can find the original data before merging [here](https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Historic-/833y-fsy8/about_data). This data is from NYCOpenData, consists some basic information: occur time, coordinate of each incident, victim's information, .etc. 

### data for additional information

#### NTA shape data

Neighborhood Tabulation Areas (NTAs) are medium-sized statistical geographies for reporting Decennial Census and American Community Survey. NTAs were delineated with the need for both geographic specificity and statistical reliability in mind. Shapefile of NTA is get from NYC Planning website, you can find the data [here](https://www.nyc.gov/site/planning/data-maps/open-data/census-download-metadata.page), or you can [download](https://s-media.nyc.gov/agencies/dcp/assets/files/zip/data-tools/bytes/nynta2020_24d.zip)

#### CDTA data

The Department of City Planning (DCP) created Community District Tabulation Areas (CDTAs) to closely approximate the 59 Community Districts of New York City for the purpose of reporting American Community Survey (ACS) data. You can find the data [here](https://www.nyc.gov/site/planning/data-maps/open-data/census-download-metadata.page), or you can [download](https://s-media.nyc.gov/agencies/dcp/assets/files/zip/data-tools/bytes/nycdta2020_24d.zip)

#### Borough Boundaries data

The Borough Boundaries dataset from the NYC Open Data portal provides detailed information about the geographical boundaries of the five boroughs of New York City: Bronx, Brooklyn, Manhattan, Queens, and Staten Island. You can find the data [here](https://data.cityofnewyork.us/City-Government/Borough-Boundaries/tqmj-j8zm)

#### Neighborhood data

[Here](https://data.insideairbnb.com/united-states/ny/new-york-city/2024-09-04/visualisations/neighbourhoods.csv) is neighbourhood list for geo filter. Sourced from city or open source GIS files. [Here](https://data.insideairbnb.com/united-states/ny/new-york-city/2024-09-04/visualisations/neighbourhoods.geojson) is GeoJSON file of neighbourhoods of the city.

#### Education data

The data source for education is shown [here](https://a816-dohbesp.nyc.gov/IndicatorPublic/data-explorer/social-conditions/?id=2334#display=summary).

#### Poverty data

The data source for poverty is [here](https://a816-dohbesp.nyc.gov/IndicatorPublic/data-explorer/economic-conditions/?id=103#display=summary)


## Data Preprocessing

The original [data](https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Historic-/833y-fsy8/about_data) only consists of some basic information and doesn’t show NTA, CDTA, Borough and neighborhood information of each incident, so we added these information to the original data base on the coordinate of each incident, we also add some demographic information: population, education level, poverty level, .etc to the data.

Here's how we added the information based on all the data we have:

* enrich NYC shooting incident data with neighborhood-level information by performing spatial operations, converts the shooting data into a spatial format, joins it with neighborhood boundaries, and adds additional neighborhood attributes for deeper spatial analysis.

* enrich NYC shooting incident data with neighborhood-level information by performing spatial operations, converts the shooting data into a spatial format, joins it with neighborhood boundaries (both 2020 and 2010 shapefiles), and adds additional neighborhood attributes for deeper spatial analysis.

* using the `holidayNYSE()` function to determine which dates are public holidays

* calculates dawn and dusk times for each unique date in the dataset. Using NYC's coordinates (`latitude` and `longitude`), the `getSunlightTimes()` function generates dawn and dusk times for each day

* reads in population data from an Excel file, filters relevant rows, selects specific columns, and merges this population data with the original dataset. 

* filters the poverty and education data for the relevant time period (2017-21) and geographic type (`NTA2020`), selects specific columns, and merges this poverty data with the shooting incident dataset. 

* calculates shooting incident rates for both Neighborhood Tabulation Areas (NTAs) and boroughs by year.





### adding neighborhood information to the shooting dataset
```{r, message=FALSE, warning=FALSE}
df=read_csv("./data/NYPD_Shooting_Incident_Data__Historic__20241119.csv")

# Ensure coordinate columns are numeric
df = df %>%
  mutate(
    X_COORD_CD = as.numeric(X_COORD_CD),
    Y_COORD_CD = as.numeric(Y_COORD_CD)
  )

# Convert to sf object with CRS EPSG:2263
df_sf = st_as_sf(df, coords = c("X_COORD_CD", "Y_COORD_CD"), crs = 2263)

# Download and read the GeoJSON file
download.file(
  url = "https://data.insideairbnb.com/united-states/ny/new-york-city/2024-09-04/visualisations/neighbourhoods.geojson",
  destfile = "neighbourhoods.geojson",
  mode = "wb"
)
neighbourhoods = st_read("neighbourhoods.geojson")

# Check and transform CRS of neighborhoods
if (st_crs(neighbourhoods)$epsg != 2263) {
  neighbourhoods = st_transform(neighbourhoods, crs = 2263)
}

# Perform the spatial join
df_joined = st_join(df_sf, neighbourhoods, left = TRUE)

# Add neighborhood information to the data frame
df$Neighborhood = df_joined$neighbourhood

# Download and use the neighborhoods CSV file
download.file(
  url = "https://data.insideairbnb.com/united-states/ny/new-york-city/2024-09-04/visualisations/neighbourhoods.csv",
  destfile = "neighbourhoods.csv",
  mode = "wb"
)
neighborhoods_csv = read_csv("neighbourhoods.csv")

# Merge additional attributes if necessary
df = df %>%
  left_join(neighborhoods_csv, by = c("Neighborhood" = "neighbourhood"))

```

converts the shooting data into a spatial format, joins it with neighborhood boundaries, and adds additional neighborhood attributes for deeper spatial analysis.

### add nta data
```{r}
df$X_COORD_CD <- as.numeric(df$X_COORD_CD)
df$Y_COORD_CD <- as.numeric(df$Y_COORD_CD)

# Convert to sf object
df_sf <- st_as_sf(df, coords = c("X_COORD_CD", "Y_COORD_CD"), crs = 2263)

# Read neighborhood shapefile
nta <- st_read("./data/nynta2020_24d/nynta2020.shp")

# Transform CRS if necessary
nta <- st_transform(nta, crs = st_crs(df_sf))

# Perform spatial join
df_joined <- st_join(df_sf, nta, left = TRUE)

# Add neighborhood information to your original data frame
df$NTA <- df_joined$NTAName  # Adjust based on actual column name
```

### add 2010 nta data
```{r}
df$X_COORD_CD <- as.numeric(df$X_COORD_CD)
df$Y_COORD_CD <- as.numeric(df$Y_COORD_CD)

# Convert to sf object
df_sf <- st_as_sf(df, coords = c("X_COORD_CD", "Y_COORD_CD"), crs = 2263)

# Read neighborhood shapefile
neighborhoods_2010 <- st_read("./data/nynta2010_24d/nynta2010.shp")

# Transform CRS if necessary
neighborhoods_2010 <- st_transform(neighborhoods_2010, crs = st_crs(df_sf))

# Perform spatial join
df_joined <- st_join(df_sf, neighborhoods_2010, left = TRUE)

# Add neighborhood information to your original data frame
df$NTA_2010 <- df_joined$NTAName  # Adjust based on actual column name
```

converts the shooting data into a spatial format, joins it with neighborhood boundaries (both 2020 and 2010 shapefiles), and adds additional neighborhood attributes for deeper spatial analysis. 


### add time information
for future study, i need to add variables related to date and time of the gunshot incidence.

#### decide whether the date of shooting is national holiday
```{r}
df$OCCUR_DATE <- as.Date(df$OCCUR_DATE, format = "%m/%d/%Y")

# Define the range of years, handling NA values
years <- seq(
  year(min(df$OCCUR_DATE, na.rm = TRUE)),
  year(max(df$OCCUR_DATE, na.rm = TRUE)),
  by = 1
)
us_holidays = holidayNYSE(years)
us_holidays = as.Date(us_holidays)
# Add a new column indicating whether the date is a holiday, year and month information of a date
df = df %>%
  mutate(
    Is_Holiday = OCCUR_DATE %in% us_holidays,
    Year = year(OCCUR_DATE),
    Month = month(OCCUR_DATE)
  )

```

The code handles holidays by using the `holidayNYSE()` function to determine which dates are public holidays. A new column (`Is_Holiday`) is added to indicate if each incident occurred on a holiday. The dataset is also enriched by extracting year and month information from the `OCCUR_DATETIME` to facilitate temporal trend analysis.

#### decide whether the sky is dark at the shooting time, using the dawn and dusk time:
```{r}
df$OCCUR_DATE <- as.character(df$OCCUR_DATE)
df$OCCUR_TIME <- as.character(df$OCCUR_TIME)

# Combine OCCUR_DATE and OCCUR_TIME into a single datetime string
datetime_str <- paste(df$OCCUR_DATE, df$OCCUR_TIME)

# Parse datetime using lubridate
df$OCCUR_DATETIME <- ymd_hms(datetime_str, tz = "America/New_York")
# Extract date from OCCUR_DATETIME
df$DATE <- as.Date(df$OCCUR_DATETIME)

# Get unique dates
unique_dates <- unique(df$DATE)
latitude <- 40.7128
longitude <- -74.0060
# Calculate dawn and dusk times
sun_times <- getSunlightTimes(
  date = unique_dates,
  lat = latitude,
  lon = longitude,
  keep = c("dawn", "dusk"),
  tz = "America/New_York"
)

# Merge with the original dataframe
df <- df %>%
  left_join(sun_times[, c("date", "dawn", "dusk")], by = c("DATE" = "date"))

# Determine if the sky is dark considering twilight
df <- df %>%
  mutate(
    Sky_Is_Dark = OCCUR_DATETIME < dawn | OCCUR_DATETIME >= dusk
  )
df <- df %>%
  select(-DATE, -dawn, -dusk)
```

The code calculates dawn and dusk times for each unique date in the dataset. Using NYC's coordinates (`latitude` and `longitude`), the `getSunlightTimes()` function generates dawn and dusk times for each day, which are then merged with the original dataset. A new column (`Sky_Is_Dark`) is added to indicate whether the shooting incident occurred during dark conditions, defined as before dawn or after dusk.


### adding total population data of each NTA

```{r}
popu <- suppressWarnings(
  read_excel("./data/nyc_detailed-race-and-ethnicity-data_2020_core-geographies.xlsx", 
             sheet = 1,
             range = "A4:I2599",
             col_types = c("numeric", "text", "text", "text", "numeric", "text", "text", "numeric", "numeric")) |>
    filter(`Orig Order` >= 2334 & `Orig Order` <= 2595) |>
    select(GeoName, NTAType, Pop) |>
    rename(Total_population = Pop)
)

df <- df |>
  left_join(popu, by = c("NTA" = "GeoName")) |>
  mutate(
    NTAType = case_when(
      NTAType == 0 ~ "Residential",
      NTAType == 9 ~ "Park",
      NTAType == 8 ~ "Airport",
      NTAType == 7 ~ "Cemetery",
      NTAType == 6 ~ "Other Special Areas",
      NTAType == 5 ~ "Rikers Island",
      is.na(NTAType) ~ "Unknown")
  )

```

reads population data from an Excel file (`nyc_detailed-race-and-ethnicity-data_2020_core-geographies.xlsx`). The data is filtered to include relevant rows for specific neighborhoods and selects columns such as `GeoName`, `NTAType`, and `Pop`. After renaming the `Pop` column to `Total_population`, this data is merged with the original dataset (`df`). The `NTAType` values are mapped to descriptive names like "Residential", "Park", or "Airport" to better understand the types of neighborhoods involved. 


### adding cdta data

```{r,message=FALSE}
df$X_COORD_CD <- as.numeric(df$X_COORD_CD)
df$Y_COORD_CD <- as.numeric(df$Y_COORD_CD)

df_sf <- st_as_sf(df, coords = c("X_COORD_CD", "Y_COORD_CD"), crs = 2263)

cdta <- st_read("./data/nycdta2020_24d/nycdta2020.shp")

cdta <- st_transform(cdta, crs = st_crs(df_sf))

df_joined <- st_join(df_sf, cdta, left = TRUE)

df$CDTA <- df_joined$CDTA2020

df$CDTA <- gsub("([A-Z]{2})(\\d{2})", "\\1 \\2", df$CDTA)

```

Similar to adding NTA data

### add poverty information
```{r}
neighborhood_poverty <- read.csv("./data/neighborhood_poverty.csv")

# Select specific columns from neighborhood_poverty
neighborhood_poverty_selected <- neighborhood_poverty %>% 
  filter(TimePeriod == '2017-21') %>% 
  filter(GeoType == 'NTA2020' ) %>% 
  select(Number, Percent, Geography)
# Merge the dataframes on 'NTA' and 'Geography', keeping all information from data_processing
df_poverty <- left_join(df, neighborhood_poverty_selected, by = c("NTA" = "Geography"))

# Rename columns after merging
df_poverty <- df_poverty %>% 
  rename(Number_poverty = Number, Percent_poverty = Percent)
```

filters the poverty data for the relevant time period (2017-21) and geographic type (`NTA2020`), selects specific columns, and merges this poverty data with the shooting incident dataset. 


### add education information
```{r}
neighborhood_education=read.csv("./data/graduated_high_school.csv")
neighborhood_education_selected <- neighborhood_education %>% 
  filter(TimePeriod == '2017-21') %>% 
  filter(GeoType == 'NTA2020' ) %>% 
  select(Number, Percent, Geography)
# Merge the dataframes on 'NTA' and 'Geography', keeping all information from data_processing
df_education <- left_join(df_poverty, neighborhood_education_selected, by = c("NTA" = "Geography"))

# Rename columns after merging
df_education <- df_education %>% 
  rename(Number_education = Number, Percent_education = Percent)
# filter between 2017 and 2023
df_education$OCCUR_DATE <- as.Date(df_education$OCCUR_DATE, format = "%Y-%m-%d")

# Filter the incidents that happened between 2017 and 2023
df_education <- df_education %>% 
  filter(OCCUR_DATE >= as.Date("2017-01-01") & OCCUR_DATE <= as.Date("2023-12-31"))
```

filters the education data for the relevant time period (2017-21) and geographic type (`NTA2020`), selects specific columns, and merges this poverty data with the shooting incident dataset. 

### compute incident rate

```{r,message=FALSE}
#add a column that shows the shooting incident rate of NTAs in that year
df_education <- df_education %>%
  group_by(NTA, Year) %>%
  mutate(incident_rate_by_year_nta = (n() / Total_population)*100) %>%
  ungroup()

#add a column that shows the shooting incident rate of boroughs in that year
boro_population <- df_education %>%
  group_by(BORO, Year) %>%
  summarise(total_population_boro = sum(Total_population, na.rm = TRUE)) %>%
  ungroup()

df_education <- df_education %>%
  left_join(boro_population, by = c("BORO", "Year")) %>%
  group_by(BORO, Year) %>%
  mutate(incident_rate_by_year_boro = (n() / total_population_boro)*100) %>%
  ungroup()

#Rename a column
df_education <- df_education %>%
  rename(Total_population_nta = Total_population)

write.csv(df_education, "data_final.csv", row.names = FALSE)
```

calculates shooting incident rates for both Neighborhood Tabulation Areas (NTAs) and boroughs by year. For each NTA, the incident rate is calculated as the number of incidents per total population (`incident_rate_by_year_nta`). Similarly, the total population for each borough is calculated, and the shooting incident rate for each borough is computed (`incident_rate_by_year_boro`). 

