---
title: "Project Report"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: sandstone
    code_folding: hide
---

```{r}
library(tidyverse)
library(plotly)
library(shiny)
library(flexdashboard)
library(viridis)
```

```{r}
data_clean <- read.csv("data_final.csv")
```


**Project Name:** *Examining Demographic Patterns in NYC Shooting Data*

**Project Members:** Chenhui Yan (CY2772), Zhaokun Lin (ZL3544), Mingyin Wang (AW3693), Zebang Zhang (ZZ3309)

## Topic

In this exploratory study, we aim to analyze the geographic and temporal distribution of shootings in New York City, with an emphasis on how socioeconomic factors influence these patterns.

## Motivation

Understanding the dynamics of shooting incidents in New York City is crucial for enhancing public safety and fostering resilient communities. By examining demographic patterns in shooting crime data alongside datasets on high school graduation rates and poverty levels, our project aims to identify the socioeconomic factors that influence gun violence. Analyzing how disparities in education and income relate to shooting incidents provides valuable insights for targeted interventions.

## Introduction

Gun violence is a significant public health and social issue New York City, where diverse communities and complex socioeconomic conditions create a unique landscape for understanding its contributing factors. This project examines the geographic, temporal, and socioeconomic patterns of shootings in NYC to identify critical influences on gun violence. We explore how shooting incidents are distributed across boroughs and neighborhoods and examine temporal trends, such as seasonal, monthly, and time-of-day variations. Additionally, we investigate the relationship between education and gun violence, focusing on whether lower high school graduation rates correlate with higher shooting prevalence. Finally, we analyze the connection between poverty and gun violence to determine if neighborhoods with higher poverty levels experience increased incidents of shootings.

## Significance

This research will aid in informing targeted public health and safety interventions by identifying the communities most impacted by shootings. Understanding these patterns will enable policymakers to implement more effective violence prevention measures and allocate resources where they are needed the most.

## Research Questions

Our project focuses on examining the socioeconomic and temporal factors that influence gun violence in New York City. Specifically, we aim to answer the following questions:

 **Geographic and Temporal Patterns of Shootings in NYC**:
   - How are shooting incidents distributed geographically across different boroughs and neighborhoods in New York City?
   - What are the temporal patterns of shootings (e.g., month, season, time of day) in different areas of NYC?

 **Education and Gun Violence**:
   - Is there an association between high school graduation rates and the prevalence of shootings across NYC neighborhoods?
   - Are neighborhoods with lower high school graduation rates more likely to experience higher rates of gun violence?

 **Poverty and Gun Violence**:
   - How does the percentage of people living below the poverty line relate to shooting incidents across neighborhoods in NYC?
   - Are higher poverty levels associated with an increased frequency of shootings?


## Data Sources and Cleaning

In this section, describe the data sources, cleaning process, and the methods used to explore the research questions.

### Data source

#### Original data

This report analyzes shooting incidents in NYC. You can find the original data before merging [here](https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Historic-/833y-fsy8/about_data). This data is from NYCOpenData, consists some basic information: occur time, coordinate of each incident, victim's information, .etc. 

#### data for additional information

##### NTA shape data

Neighborhood Tabulation Areas (NTAs) are medium-sized statistical geographies for reporting Decennial Census and American Community Survey. NTAs were delineated with the need for both geographic specificity and statistical reliability in mind. Shapefile of NTA is get from NYC Planning website, you can find the data [here](https://www.nyc.gov/site/planning/data-maps/open-data/census-download-metadata.page), or you can [download](https://s-media.nyc.gov/agencies/dcp/assets/files/zip/data-tools/bytes/nynta2020_24d.zip)

##### CDTA data

The Department of City Planning (DCP) created Community District Tabulation Areas (CDTAs) to closely approximate the 59 Community Districts of New York City for the purpose of reporting American Community Survey (ACS) data. You can find the data [here](https://www.nyc.gov/site/planning/data-maps/open-data/census-download-metadata.page), or you can [download](https://s-media.nyc.gov/agencies/dcp/assets/files/zip/data-tools/bytes/nycdta2020_24d.zip)

##### Borough Boundaries data

The Borough Boundaries dataset from the NYC Open Data portal provides detailed information about the geographical boundaries of the five boroughs of New York City: Bronx, Brooklyn, Manhattan, Queens, and Staten Island. You can find the data [here](https://data.cityofnewyork.us/City-Government/Borough-Boundaries/tqmj-j8zm)

##### Neighborhood data

[Here](https://data.insideairbnb.com/united-states/ny/new-york-city/2024-09-04/visualisations/neighbourhoods.csv) is neighbourhood list for geo filter. Sourced from city or open source GIS files. [Here](https://data.insideairbnb.com/united-states/ny/new-york-city/2024-09-04/visualisations/neighbourhoods.geojson) is GeoJSON file of neighbourhoods of the city.

##### Education data

The data source for education is shown [here](https://a816-dohbesp.nyc.gov/IndicatorPublic/data-explorer/social-conditions/?id=2334#display=summary).

##### Poverty data

The data source for poverty is [here](https://a816-dohbesp.nyc.gov/IndicatorPublic/data-explorer/economic-conditions/?id=103#display=summary)


### Data processing

The original [data](https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Historic-/833y-fsy8/about_data) only consists of some basic information and doesn’t show NTA, CDTA, Borough and neighborhood information of each incident, so we added these information to the original data base on the coordinate of each incident, we also add some demographic information: population, education level, poverty level, .etc to the data.

Here's how we added the information based on all the data we have:

* enrich NYC shooting incident data with neighborhood-level information by performing spatial operations, converts the shooting data into a spatial format, joins it with neighborhood boundaries, and adds additional neighborhood attributes for deeper spatial analysis.

* enrich NYC shooting incident data with neighborhood-level information by performing spatial operations, converts the shooting data into a spatial format, joins it with neighborhood boundaries (both 2020 and 2010 shapefiles), and adds additional neighborhood attributes for deeper spatial analysis.

* using the `holidayNYSE()` function to determine which dates are public holidays

* calculates dawn and dusk times for each unique date in the dataset. Using NYC's coordinates (`latitude` and `longitude`), the `getSunlightTimes()` function generates dawn and dusk times for each day

* reads in population data from an Excel file, filters relevant rows, selects specific columns, and merges this population data with the original dataset. 

* filters the poverty and education data for the relevant time period (2017-21) and geographic type (`NTA2020`), selects specific columns, and merges this poverty data with the shooting incident dataset. 

* calculates shooting incident rates for both Neighborhood Tabulation Areas (NTAs) and boroughs by year.

The detailed code is [here](data_processing.html)

### Neighborhood Poverty
```{r}
neighborhood_poverty <- read.csv("./data/neighborhood_poverty.csv")

# Select specific columns from neighborhood_poverty
neighborhood_poverty_selected <- neighborhood_poverty %>% 
  filter(TimePeriod == '2017-21') %>% 
  filter(GeoType == 'NTA2020' ) %>% 
  select(Number, Percent, Geography)

```

* Downloaded the full table from the NYC Environment and Health Data Portal
* Filtered to include only percent data from 2017-2021 for each NTA 
* Describes estimated percentage of people whose annual income falls below 100% of the federal poverty level in each NTA

### Graduated High School
```{r}
neighborhood_education=read.csv("./data/graduated_high_school.csv")
neighborhood_education_selected <- neighborhood_education %>% 
  filter(TimePeriod == '2017-21') %>% 
  filter(GeoType == 'NTA2020' ) %>% 
  select(Number, Percent, Geography)

```

* Downloaded the full table from the NYC Environment and Health Data Portal
* Filtered the data to include only the percentage information for high school graduation between 2017 and 2021, focusing on Neighborhood Tabulation Areas (NTAs) as defined by the 2020 census.
* This dataset describes the estimated percentage of individuals aged 25 and older who completed high school or obtained a high school equivalency in each NTA

## Exploratory data analysis
```{r}

```

## Maps
```{r}

```

## Statistical analysis
## Method for Calculating Correlation Coefficients
To assess potential linear relationships between socioeconomic variables and shooting rates, correlation coefficients were calculated. Specifically, the relationship between poverty rates and shooting rates, as well as between the percentage of high school graduates and shooting rates, was explored across different neighborhoods in NYC.

### Poverty

Calculate the correlation between the poverty percentage and the incident rate.
```{r}
correlation <- cor(data_clean$incident_rate_by_year_nta, data_clean$Percent_poverty, use = "complete.obs")
print(paste("Correlation coefficient: ", correlation))
```

```{r}

data_clean %>%
  plot_ly(x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA, colors = "viridis", 
          type = "scatter", mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", BORO, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) %>%
  layout(title = "Percent Below the Poverty Line and Incident Rate in NYC",
         xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
         yaxis = list(title = 'Incident Rate'),
         legend = list(title = list(text = 'Neighborhood')))
```
```{r}
# Scatter plot for Brooklyn
data_clean |> 
  filter(neighbourhood_group == "Brooklyn") |> 
  plot_ly(data = _, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "plasma", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |> 
    layout(title = "Percent Below the Poverty Line and Incident Rate in Brooklyn",
           xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))

```

```{r}
# Scatter plot for Staten Island
data_clean |> 
  filter(neighbourhood_group == "Staten Island") |> 
  plot_ly(data = _, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "inferno", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |> 
    layout(title = "Percent Below the Poverty Line and Incident Rate in Staten Island",
           xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))

```

* Across all neighborhoods in NYC, there is a moderate positive linear relationship between the percentage of people below the poverty line (Percent_poverty) and the incident rate by neighborhood (incident_rate_by_year_nta).(r = 0.508).
* Associations differ by borough.
* Notably, Brooklyn (0.5686) and Staten Island (0.5576) show the highest correlations, suggesting that the relationship between poverty and incident rate is stronger in these boroughs.

### Education

Calculate the correlation between the graduated in highschool percentage and the incident rate
```{r}
# Calculate the correlation between the graduated in highschool percentage and the incident rate
correlation <- cor(data_clean$incident_rate_by_year_nta, data_clean$Percent_education, use = "complete.obs")
print(paste("Correlation coefficient: ", correlation))

```

```{r}
# Create a scatter plot to visualize the relationship
data_clean %>%
  plot_ly(x = ~Percent_education, y = ~incident_rate_by_year_nta, 
          color = ~NTA, colors = "viridis", 
          type = "scatter", mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", BORO, 
                        "<br>% graduated HS: ", Percent_education, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) %>%
  layout(title = "Percent graduated high school and Incident Rate in NYC",
         xaxis = list(title = 'Percentage of People graduated in high school'),
         yaxis = list(title = 'Incident Rate'),
         legend = list(title = list(text = 'Neighborhood')))
```

```{r}
# Scatter plot for The Bronx
data_clean |> 
  filter(neighbourhood_group == "Bronx") |>
  plot_ly(data = _, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "magma", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% graduated HS: ", Percent_education, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |> 
   layout(title = "Percent graduated high school and Incident Rate in The Bronx",
           xaxis = list(title = 'Percentage of People graduated in high school'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))


```

```{r}
# Scatter plot for Staten Island
data_clean |> 
  filter(neighbourhood_group == "Staten Island") |>
  plot_ly(data = _, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "inferno", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% graduated HS: ", Percent_education, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |> 
   layout(title = "Percent graduated high school and Incident Rate in Staten Island",
           xaxis = list(title = 'Percentage of People graduated in Staten Island'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))

```

The correlation coefficient is -0.2748, indicating a negative relationship between the percentage of people who graduated from high school and the incident rate. This means that there is a tendency for higher education levels to be associated with lower incident rates
* Across all neighborhoods in NYC,there is no clear linear trend between high school graduation rates and incident rates in Manhattan.
* Associations differ by borough.
* Notably, The Bronx (-0.3996) and Staten Island (-0.6452) show the highest correlations, suggesting that the relationship between percentage of people who graduated from high school and incident rate is stronger in these boroughs.

## Discussion

## Limitations

