---
title: "Examining Demographic Patterns in NYC Shooting Data"
output: 
  html_document:
    css: bootstrap.min.css
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,  
  message = FALSE   
)
```

Welcome to the NYC Shooting Incident Report website. This platform provides a comprehensive [analysis](Final_Report.html) of shooting incidents across New York City from 2017 to 2023. Using data-driven insights, we explore [geographic, periodic](descriptive_statistics.html) and [social-economic patterns](Poverty_Education.html) behind these incidents, aiming to enhance public understanding and support policy-making for a safer city. Navigate through our interactive visualizations to understand how factors like neighborhood characteristics, poverty rates, and time of day influence shooting trends across NYC.



## See all shooting incident in NYC (2017-2023) on a map!

Hover over each point on the map to learn more about the coordinate and NTA(Neighborhood Tabulation Areas) of each incident, you can also know the time and whether the sky is dark at that incident time.

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(plotly)

# 1. Load your incidents data
incidents_data <- read.csv("data_final.csv")

# 2. Filter data from 2017 to 2023 and remove rows with missing values in key columns
incidents_filtered <- incidents_data %>%
  filter(Year >= 2017 & Year <= 2023) %>%
  drop_na(Sky_Is_Dark, Longitude, Latitude, NTA, OCCUR_TIME)  # Ensure essential columns have no NA


# Optional: Relabel Sky_Is_Dark for better readability
incidents_filtered <- incidents_filtered %>%
  mutate(Sky_Condition = ifelse(Sky_Is_Dark, "Night", "Day"))

# 3. Create the interactive map using the filtered data
plot_ly(data = incidents_filtered, 
        type = "scattermapbox",       # Specifies that we're creating a Mapbox scatter plot
        mode = "markers",             # Use markers to represent data points
        lon = ~Longitude,             # Longitude for x-axis (map)
        lat = ~Latitude,              # Latitude for y-axis (map)
        color = ~Sky_Condition,       # Color markers based on Sky_Condition
        colors = c("Day" = "orange", "Night" = "blue"), # Assign specific colors
        marker = list(size = 5,        # Set marker size
                      opacity = 0.7),  # Set marker opacity for better visibility
        text = ~paste("NTA:", NTA,
                      "<br>Time of Incident:", OCCUR_TIME,
                      "<br>Sky Condition:", Sky_Condition)  # Hover text
) |> 
  layout(
    title = "DAY OR NIGHT Incidents in NYC",  # Set the plot title
    mapbox = list(
      style = "carto-positron",      # Map style
      center = list(lon = -73.92, lat = 40.70),  # Center the map on NYC
      zoom = 9.5                      # Set zoom level
    ),
    showlegend = TRUE,               # Display the legend
    legend = list(title = list(text = 'Sky Condition'))  # Legend title
  )

```

## Explore more trends:

* see the geographic and periodic [distribution](descriptive_statistics.html)!

* see [association](https://zz3309.shinyapps.io/poverty_education/) between incident rate and economic-education condition!

* see some more interactive [plots](https://zz3309.shinyapps.io/nyc_shooting_incident_shiny_dashboard/)!

## Screen cast 

<iframe width="560" height="315" src="https://www.youtube.com/embed/6y5CqAHxGX0" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>