---
title: "Poverty VS Incidence Rate"
author: "Chenhui Yan"
date: "2024-12-02"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,  
  message = FALSE   
)
```

```{r}
library(tidyverse)
library(plotly)
library(shiny)
library(flexdashboard)
library(viridis)
```

## Poverty

Is there an association between the percentage of people whose income is below the poverty line and the incidence rate 1) across neighborhoods in NYC and 2) across neighborhoods in each borough?

```{r}
data_clean <- read.csv("data_final.csv")
```

## Poverty across neighborhoods in NYC

Calculate the correlation between the poverty percentage and the incident rate.
```{r}
correlation <- cor(data_clean$incident_rate_by_year_nta, data_clean$Percent_poverty, use = "complete.obs")
print(paste("Correlation coefficient: ", correlation))
```

When examining all of the neighborhoods in NYC, there is a moderate positive linear relationship between the percentage of people below the poverty line (Percent_poverty) and the incident rate by neighborhood (incident_rate_by_year_nta).(r = 0.508).

```{r}
data_clean %>%
  plot_ly(x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA, colors = "viridis", 
          type = "scatter", mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", BORO, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) %>%
  layout(title = "Percent Below the Poverty Line and Incident Rate in NYC",
         xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
         yaxis = list(title = 'Incident Rate'),
         legend = list(title = list(text = 'Neighborhood')))
```

This plot suggests that neighborhoods with higher poverty rates tend to have higher incident rates.

## Poverty By Borough

```{r}
selectInput(
  "borough_choice", 
  label = h3("Borough"),  
  choices = c("Manhattan","Brooklyn","Queens","Bronx","Staten Island"), selected = "Manhattan")
```

```{r}
compute_correlation <- function(data,borough_name) {
  correlation <- cor(
    data$incident_rate_by_year_nta,
    data$Percent_poverty,
    use = "complete.obs"
  )
  paste("Correlation coefficient for", borough_name, ":", correlation)
}

renderPrint({
  filtered_data=data_clean %>%
    filter(neighbourhood_group == input[["borough_choice"]])
  compute_correlation(filtered_data,input[["borough_choice"]])
})
```

* Positive Correlation Across All Boroughs:
 * The correlation coefficients for all boroughs are positive, indicating that as the percentage of people below the poverty line increases, the incident rate also tends to increase.
* Differences in Correlation Strength:
 * Brooklyn (0.5686) and Staten Island (0.5576) show the highest correlations, suggesting that the relationship between poverty and incident rate is stronger in these boroughs.
 * Manhattan (0.3345) and Queens (0.2797) show weaker correlations, indicating that the relationship between poverty and incident rate is less pronounced in these areas.
 * Bronx (0.4463) has a moderate correlation, which also points to a positive but not as strong relationship.

```{r}
renderPlotly({
  filtered_data=data_clean %>%
    filter(neighbourhood_group == input[["borough_choice"]])
  plot_ly(data = filtered_data, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "viridis", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |>
     layout(title = paste("Percent Below the Poverty Line and Incident Rate in", input[["borough_choice"]]),
           xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))
})
```

Some neighborhoods, like East Harlem (North), East Harlem (South), and Chinatown-Two Bridges, seem to have both higher poverty percentages and higher incident rates, suggesting a more direct connection between poverty and incidents.

Brownsville, which has a high poverty rate (between 30-50%), also has some of the highest incident rates in Brooklyn, indicating that poverty may significantly impact incident rates in this neighborhood.

Claremont Village-Claremont (East) and Fordham Heights seem to have higher incident rates compared to other neighborhoods with similar poverty levels

Corona and East Elmhurst show relatively high incident rates, particularly compared to neighborhoods with similar poverty percentages. 

In Staten Island, the relationship between poverty and incident rates is less pronounced compared to other boroughs. 

```{r}
incidents_data <- read.csv("data_final.csv")

incidents_filtered <- incidents_data %>%
  filter(Year >= 2017 & Year <= 2023)

#Aggregate total incidents by CDTA
incidents_by_cdta <- incidents_filtered %>%
  group_by(CDTA) %>%
  summarise(
    total_incidents = n(),
    Longitude = mean(Longitude, na.rm = TRUE),
    Latitude = mean(Latitude, na.rm = TRUE)
  ) %>%
  ungroup()

# Define initial breaks from 0 to 400 with intervals of 5
initial_breaks <- seq(0, 400, by = 5)

# Calculate the maximum number of incidents
max_incidents <- max(incidents_by_cdta$total_incidents, na.rm = TRUE)

# Check if maximum incidents exceed 400 and adjust breaks and labels accordingly
if (max_incidents > 400) {
  # Extend breaks to include all incidents > 400
  extended_breaks <- c(initial_breaks, Inf)
  
  # Create labels for all intervals, including the last one for incidents >400
  labels <- c(
    paste(initial_breaks[-length(initial_breaks)], initial_breaks[-1], sep = "-"),
    paste0("401-", max_incidents)
  )
} else {
  # No need to extend breaks
  extended_breaks <- initial_breaks
  labels <- paste(initial_breaks[-length(initial_breaks)], initial_breaks[-1], sep = "-")
}

# Validate that the number of labels matches the number of intervals
num_intervals <- length(extended_breaks) - 1
num_labels <- length(labels)

if (num_intervals != num_labels) {
  stop("Number of labels does not match the number of intervals.")
}

# Assign incident ranges
incidents_by_cdta <- incidents_by_cdta %>%
  mutate(
    incident_range = cut(
      total_incidents,
      breaks = extended_breaks,
      include.lowest = TRUE,
      right = FALSE,  # Left-inclusive intervals [a, b)
      labels = labels
    )
  )

# Verify the distribution of incident ranges
print(table(incidents_by_cdta$incident_range, useNA = "ifany"))

# Check for missing coordinates and exclude them
missing_coords <- incidents_by_cdta %>%
  filter(is.na(Longitude) | is.na(Latitude))

if (nrow(missing_coords) > 0) {
  cat("Warning: The following CDTAs have missing coordinates and will be excluded from the plot:\n")
  print(missing_coords$CDTA)
  
  # Exclude CDTAs with missing coordinates
  incidents_by_cdta <- incidents_by_cdta %>%
    filter(!is.na(Longitude) & !is.na(Latitude))
}

Sys.setenv('MAPBOX_TOKEN' = 'pk.eyJ1IjoiZWR3YXJkenp6IiwiYSI6ImNtNDkwMzZ3YTA1MGIyanE1ZjdlNTZodjcifQ.BsLSffH3a0z5Xqf1Tqc4ag')  

# Create the scatter map
incident_map <- plot_ly(
  data = incidents_by_cdta,
  type = "scattermapbox",
  mode = "markers",
  lon = ~Longitude,  
  lat = ~Latitude,   
  color = ~incident_range,
  colors = "viridis",  
  marker = list(
    size = 8,
    opacity = 0.7
  ),
  text = ~paste(
    "CDTA: ", CDTA,
    "<br>Total Incidents: ", total_incidents,
    "<br>Incident Range: ", incident_range
  )
) %>%
  layout(
    title = "Total Number of Incidents Across NYC CDTAs (2017-2023)",
    mapbox = list(
      style = "carto-positron",  # Choose desired map style
      center = list(lon = -74.0060, lat = 40.7128),  # Centered on NYC
      zoom = 10
    ),
    legend = list(title = list(text = 'Incident Range'))
  )

incident_map
```
