---
title: "Poverty/Education VS Incidence Rate"
author: "Chenhui Yan"
date: "2024-12-02"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,  
  message = FALSE   
)
```

```{r}
library(tidyverse)
library(plotly)
library(shiny)
library(flexdashboard)
library(viridis)
```

# Poverty

Is there an association between the percentage of people whose income is below the poverty line and the incidence rate 1) across neighborhoods in NYC and 2) across neighborhoods in each borough?

```{r}
data_clean <- read.csv("data_final.csv")
```

## Poverty across neighborhoods in NYC

Calculate the correlation between the poverty percentage and the incident rate.
```{r}
correlation <- cor(data_clean$incident_rate_by_year_nta, data_clean$Percent_poverty, use = "complete.obs")
print(paste("Correlation coefficient: ", correlation))
```

When examining all of the neighborhoods in NYC, there is a moderate positive linear relationship between the percentage of people below the poverty line (Percent_poverty) and the incident rate by neighborhood (incident_rate_by_year_nta).(r = 0.508).

```{r}
data_clean %>%
  plot_ly(x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA, colors = "viridis", 
          type = "scatter", mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", BORO, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) %>%
  layout(title = "Percent Below the Poverty Line and Incident Rate in NYC",
         xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
         yaxis = list(title = 'Incident Rate'),
         legend = list(title = list(text = 'Neighborhood')))
```

This plot suggests that neighborhoods with higher poverty rates tend to have higher incident rates.

## Poverty By Borough

```{r}
selectInput(
  "borough_choice", 
  label = h3("Borough"),  
  choices = c("Manhattan","Brooklyn","Queens","Bronx","Staten Island"), selected = "Manhattan")
```

```{r}
compute_correlation_po <- function(data,borough_name) {
  correlation <- cor(
    data$incident_rate_by_year_nta,
    data$Percent_poverty,
    use = "complete.obs"
  )
  paste("Correlation coefficient for", borough_name, ":", correlation)
}

renderPrint({
  filtered_data=data_clean %>%
    filter(neighbourhood_group == input[["borough_choice"]])
  compute_correlation_po(filtered_data,input[["borough_choice"]])
})
```

* Positive Correlation Across All Boroughs:
 * The correlation coefficients for all boroughs are positive, indicating that as the percentage of people below the poverty line increases, the incident rate also tends to increase.
* Differences in Correlation Strength:
 * Brooklyn (0.5686) and Staten Island (0.5576) show the highest correlations, suggesting that the relationship between poverty and incident rate is stronger in these boroughs.
 * Manhattan (0.3345) and Queens (0.2797) show weaker correlations, indicating that the relationship between poverty and incident rate is less pronounced in these areas.
 * Bronx (0.4463) has a moderate correlation, which also points to a positive but not as strong relationship.

```{r}
renderPlotly({
  filtered_data=data_clean %>%
    filter(neighbourhood_group == input[["borough_choice"]])
  plot_ly(data = filtered_data, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "viridis", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |>
     layout(title = paste("Percent Below the Poverty Line and Incident Rate in", input[["borough_choice"]]),
           xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))
})
```

Some neighborhoods, like East Harlem (North), East Harlem (South), and Chinatown-Two Bridges, seem to have both higher poverty percentages and higher incident rates, suggesting a more direct connection between poverty and incidents.

Brownsville, which has a high poverty rate (between 30-50%), also has some of the highest incident rates in Brooklyn, indicating that poverty may significantly impact incident rates in this neighborhood.

Claremont Village-Claremont (East) and Fordham Heights seem to have higher incident rates compared to other neighborhoods with similar poverty levels

Corona and East Elmhurst show relatively high incident rates, particularly compared to neighborhoods with similar poverty percentages. 

In Staten Island, the relationship between poverty and incident rates is less pronounced compared to other boroughs. 

# Education

Is there an association between the percentage of people who graduated high school and the incidence rate 1) across neighborhoods in NYC and 2) across neighborhoods in each borough?

## Education across neighborhoods in NYC

Calculate the correlation between the graduated in high school percentage and the incident rate
```{r}
correlation <- cor(data_clean$incident_rate_by_year_nta, data_clean$Percent_education, use = "complete.obs")
print(paste("Correlation coefficient: ", correlation))
```

The correlation coefficient is -0.2748, indicating a negative relationship between the percentage of people who graduated from high school and the incident rate. This means that there is a tendency for higher education levels to be associated with lower incident rates.

```{r}
data_clean %>%
  plot_ly(x = ~Percent_education, y = ~incident_rate_by_year_nta, 
          color = ~NTA, colors = "viridis", 
          type = "scatter", mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", BORO, 
                        "<br>% graduated HS: ", Percent_education, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) %>%
  layout(title = "Percent graduated high school and Incident Rate in NYC",
         xaxis = list(title = 'Percentage of People graduated in high school'),
         yaxis = list(title = 'Incident Rate'),
         legend = list(title = list(text = 'Neighborhood')))
```

For neighborhoods with a high percentage of high school graduates (90% and above), incident rates are generally low, often below 0.05. This suggests that higher educational attainment may be associated with fewer incidents.

## Education By Borough

```{r}
selectInput(
  "borough_choice", 
  label = h3("Borough"),  
  choices = c("Manhattan","Brooklyn","Queens","Bronx","Staten Island"), selected = "Manhattan")
```

```{r}
compute_correlation_edu <- function(data, borough_name) {
  correlation <- cor(
    data$incident_rate_by_year_nta,
    data$Percent_education,
    use = "complete.obs"
  )
  paste("Correlation coefficient for", borough_name, ":", correlation)
}

renderPrint({
  filtered_data=data_clean %>%
    filter(neighbourhood_group == input[["borough_choice"]])
  compute_correlation_edu(filtered_data,input[["borough_choice"]])
})
```

* Manhattan: -0.2284
 * There is a weak negative correlation between high school graduation rates and incident rates in Manhattan. This suggests that as the percentage of high school graduates increases, the incident rate tends to decrease slightly. 
 
* Brooklyn: -0.2487
 * Brooklyn also shows a weak negative correlation. The trend is similar to Manhattan, where higher graduation rates are associated with slightly lower incident rates.

* Bronx: -0.3996
 * The Bronx has a moderate negative correlation between graduation rates and incident rates. This suggests a stronger relationship compared to Manhattan and Brooklyn, indicating that in the Bronx, neighborhoods with higher percentages of high school graduates are more likely to have lower incident rates.

* Staten Island: -0.6452
 * Staten Island has a strong negative correlation between high school graduation rates and incident rates.

* Queens: 0.1007
 * Queens shows a weak positive correlation. Unlike the other boroughs, this suggests that in Queens, neighborhoods with higher graduation rates may have slightly higher incident rates, but this relationship is very weak. 
 
```{r}
renderPlotly({
  filtered_data=data_clean %>%
    filter(neighbourhood_group == input[["borough_choice"]])
  plot_ly(data = filtered_data, x = ~Percent_education, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "viridis", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_education, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |>
     layout(title = paste("Percent graduated high school and Incident Rate in", input[["borough_choice"]]),
           xaxis = list(title = 'Percentage of People graduated in high school'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))
})
```

There is no clear linear trend between high school graduation rates and incident rates in Manhattan. The distribution of points suggests a weak negative correlation overall.East Harlem (North and South) and Chinatown-Two Bridges appear to have both lower high school graduation rates and higher incident rates. This suggests that in these neighborhoods, lower educational attainment may be correlated with higher vulnerability.

Brownsville and East Flatbush are notable for having both low educational attainment and high incident rates, indicating a relationship between these two variables.

Co-op City and Morris Park are examples of neighborhoods with higher graduation rates and lower incident rates. This may suggest that neighborhoods with better educational attainment benefit from lower rates of incidents.

Far Rockaway-Bayswater and Corona tend to have both lower educational attainment and higher incident rates.

Neighborhoods with higher percentages of high school graduates (above 15%) generally have incident rates below 0.02.Port Richmond and St. George-New Brighton are notable for having higher incident rates compared to other neighborhoods with similar graduation rates.